---
title: "Case Study 2 - Employee Attrition and Salary Prediction"
author: "Puri Rudick - https://purifect.shinyapps.io/Employee_Attrition/"
date: "2/21/2021"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---
The analysis of employee attrition and salary prediction was conducted using 870 observations (employees) dataset from DDSAnalytics Company.
Some of the focuses of this exploratory data analysis were placed on attrition and salary prediction, and trends for specific job roles.

This analysis is useful for DDSAnalytics future business decision.
```{r setup, include=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(ggplot2)
library(fmsb)
library(stringr)
library(class)
library(caret)
library(corrplot)
library(Metrics)
library(Hmisc)
library(MASS)
library(GGally)
library(leaps)
library(readxl)
install.packages('e1071', dependencies=TRUE)
```

```{r}
# Read in the data
DataOrig = read.csv("CaseStudy2-data.csv",header = TRUE)

# Explore missing values in each columns - There is none!
sapply(DataOrig, function(x) sum(is.na(x)))
```

### Trends for Specific Job Roles

Job Role vs. Attrition
```{r message=FALSE, warning=FALSE}
df.JobRole = DataOrig %>% group_by(JobRole, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.JobRole %>% ggplot(aes(x=JobRole, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3) + labs(title = 'Bar Plot Job Role vs. Attrition', x = 'Job Role', y = 'Percent Attrition') +
  theme_classic() + scale_fill_brewer(palette="Blues") + coord_flip()
```

Job Role vs. Environment Satisfaction, Job Satisfaction, and Work Life Balance
```{r message=FALSE, warning=FALSE}
#Prepare data for radar plot
JobRoleData <- DataOrig %>% 
  dplyr::group_by(JobRole) %>% 
  dplyr::summarize(EnvironmentSatisfaction_mean = mean(EnvironmentSatisfaction),     
                   JobSatisfaction_mean = mean(JobSatisfaction),     
                   WorkLifeBalance_mean = mean(WorkLifeBalance))
JobRoleData <-as.data.frame(JobRoleData)

n <- JobRoleData$JobRole
JobRoleData <- as.data.frame(t(JobRoleData[,-1]))
colnames(JobRoleData) <- n
names(JobRoleData) <- sub(" ", ".", names(JobRoleData))


max_min <- data.frame(
  `Healthcare Representative` = c(3, 2.5), 
  `Human Resources` = c(3, 2.5),
  `Laboratory Technician` = c(3, 2.5),
  `Manager` = c(3, 2.5),
  `Manufacturing Director` = c(3, 2.5),
  `Research Director`= c(3, 2.5),
  `Research Scientist` = c(3, 2.5),
  `Sales Executive`= c(3, 2.5),
  `Sales Representative`= c(3, 2.5))
rownames(max_min) <- c("Max", "Min")

df <- rbind(max_min, JobRoleData)

create_beautiful_radarchart <- function(data, color = "#00AFBB", 
                                        vlabels = colnames(data), vlcex = 1,
                                        caxislabels = NULL, title = NULL, ...){
  radarchart(
    data, axistype = 1,
    # Customize the polygon
    pcol = color, pfcol = scales::alpha(color, 0.2), plwd = 2, plty = 1,
    # Customize the grid
    cglcol = "grey", cglty = 1, cglwd = 0.8,
    # Customize the axis
    axislabcol = "grey", 
    # Variable labels
    vlcex = vlcex, vlabels = vlabels,
    caxislabels = caxislabels, title = title, ...
  )
}

par(mar=c(0,0,0,0))
# Create the radar charts
create_beautiful_radarchart(
  data = df, caxislabels = c(2.5,2.6,2.7,2.8,2.9),
  color = c("#00AFBB", "#E7B800", "#FC4E07")
)
# Add an horizontal legend
legend(
  x=.75, y=-.8, legend = c("Envi Satisfaction","Job Satisfaction", "Work Life Balance"), horiz = FALSE,
  bty = "n", pch = 1 , col = c("#00AFBB", "#E7B800", "#FC4E07"),
  text.col = "black", cex = 1, pt.cex = 1.5
  )

```

### Top Three Factors for Attrition
To Identify the top three factors for attrition, the analyst created an RShiny app (in the link below) to help visualization the dataset

https://purifect.shinyapps.io/Employee_Attrition/

The top three factors responsible for attrition are Monthly Income, Job Involvement, and Overtime
```{r, include=FALSE}
# Identify the top three factors that contribute to turnover 
# Age - Quite important - Younger > Older
df.Age = DataOrig %>% group_by(Age, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.Age %>% ggplot(aes(x=Age, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)

ggplot(DataOrig, aes(x=Attrition, y=Age)) + geom_boxplot()

#JobRole vs. Age
ggplot(DataOrig, aes(x=JobRole, y=Age, color=Attrition)) +
  geom_point()

# Business travel - Quite important
df.BusTravel = DataOrig %>% group_by(BusinessTravel, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.BusTravel %>% ggplot(aes(x=BusinessTravel, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# Daily rate (quite), Hourly rate (NOT), Monthly Rate (NOT), and Monthly Income (quite)
ggplot(DataOrig, aes(x=Attrition, y=DailyRate)) + geom_boxplot() # Attrition group has less daily rate
ggplot(DataOrig, aes(x=Attrition, y=HourlyRate)) + geom_boxplot() # Attrition group has more hourly rate which does not make sense..
ggplot(DataOrig, aes(x=Attrition, y=MonthlyRate)) + geom_boxplot() # Attrition group has more monthly rate which does not make sense..
ggplot(DataOrig, aes(x=Attrition, y=MonthlyIncome)) +
  geom_boxplot(fill="lightskyblue1", color="navyblue") +
  labs(title = 'Box Plot Monthly Income vs. Attrition', x = 'Attrition', y = 'Monthly Income') +
  theme_classic() + coord_flip()
tapply(DataOrig$MonthlyIncome, DataOrig$Attrition, summary)

# Department - Quite important - Sale > R&D > HR
df.Department = DataOrig %>% group_by(Department, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.Department %>% ggplot(aes(x=Department, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# Distance from home - Quite important
df.Distance = DataOrig %>% group_by(DistanceFromHome, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.Distance %>% ggplot(aes(x=DistanceFromHome, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)

ggplot(DataOrig, aes(x=Attrition, y=DistanceFromHome)) + geom_boxplot() # The more far, the more


# Education - NOT important factor
df.Education = DataOrig %>% group_by(Education, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.Education %>% ggplot(aes(x=Education, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# Education field - NOT important factor
df.EduField = DataOrig %>% group_by(EducationField, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.EduField %>% ggplot(aes(x=EducationField, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# Environment satisfaction - NOT important factor
df.Envi = DataOrig %>% group_by(EnvironmentSatisfaction, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.Envi %>% ggplot(aes(x=EnvironmentSatisfaction, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# Gender, Male > Female - NOT important factor
df.Gender = DataOrig %>% group_by(Gender, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.Gender %>% ggplot(aes(x=Gender, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# Job involvement
df.JobInvolvement = DataOrig %>% group_by(JobInvolvement, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.JobInvolvement %>% ggplot(aes(x=JobInvolvement, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3) + labs(title = 'Bar Plot Job Involvement vs. Attrition', x = 'Job Involvement', y = 'Percent Attrition') +
  theme_classic() + scale_fill_brewer(palette="Blues")


# Job level - quite Important
df.JobLevel = DataOrig %>% group_by(JobLevel, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.JobLevel %>% ggplot(aes(x=JobLevel, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3) +
  labs(title = 'Bar Plot Job Level vs. Attrition', x = 'Job Level', y = 'Percent Attrition') +
  theme_classic() + scale_fill_brewer(palette="Blues")


# Job Satisfaction, The less satisfaction, the more chance for attrition
df.JobSatisfaction = DataOrig %>% group_by(JobSatisfaction, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.JobSatisfaction %>% ggplot(aes(x=JobSatisfaction, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# Marital Status, Single > Married > Divorced - NOT important factor
df.MaritalStatus = DataOrig %>% group_by(MaritalStatus, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.MaritalStatus %>% ggplot(aes(x=MaritalStatus, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)

# NumCompaniesWorked, 5+ companies, more attrition
df.NumCompaniesWorked = DataOrig %>% group_by(NumCompaniesWorked, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.NumCompaniesWorked %>% ggplot(aes(x=NumCompaniesWorked, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)

# Over time
df.OT = DataOrig %>% group_by(OverTime, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.OT %>% ggplot(aes(x=OverTime, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3) + labs(title = 'Bar Plot Overtime vs. Attrition', x = 'Overtime', y = 'Percent Attrition') +
  theme_classic() + scale_fill_brewer(palette="Blues")


# PercentSalaryHike, 22%-24% (from 25% max) has more chance of attribute which doesn't make sense - NOT important
df.PercentSalaryHike = DataOrig %>% group_by(PercentSalaryHike, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.PercentSalaryHike %>% ggplot(aes(x=PercentSalaryHike, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# PerformanceRatinge, not significant different between 3 and 4 - NOT important
df.PerformanceRating = DataOrig %>% group_by(PerformanceRating, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.PerformanceRating %>% ggplot(aes(x=PerformanceRating, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# RelationshipSatisfaction, for 1-3, the better relationship, the less chance for attrition but 4 has higher % attrition than 2 and 3 ?? - NOT important
df.RelationshipSatisfaction = DataOrig %>% group_by(RelationshipSatisfaction, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.RelationshipSatisfaction %>% ggplot(aes(x=RelationshipSatisfaction, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# StockOptionLevel, for 1-3, the higher stock level, the less chance for attrition but 4 has higher % attrition than 2 and 3 ?? - NOT important
df.StockOptionLevel = DataOrig %>% group_by(StockOptionLevel, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.StockOptionLevel %>% ggplot(aes(x=StockOptionLevel, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# TotalWorkingYears - NOT important
df.TotalWorkingYears = DataOrig %>% group_by(TotalWorkingYears, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.TotalWorkingYears %>% ggplot(aes(x=TotalWorkingYears, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# TrainingTimesLastYear - NOT important
df.TrainingTimesLastYear = DataOrig %>% group_by(TrainingTimesLastYear, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.TrainingTimesLastYear %>% ggplot(aes(x=TrainingTimesLastYear, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# WorkLifeBalance - quite important
df.WorkLifeBalance = DataOrig %>% group_by(WorkLifeBalance, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.WorkLifeBalance %>% ggplot(aes(x=WorkLifeBalance, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# YearsAtCompany - NOT important
df.YearsAtCompany = DataOrig %>% group_by(YearsAtCompany, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.YearsAtCompany %>% ggplot(aes(x=YearsAtCompany, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# YearsInCurrentRole - NOT important
df.YearsInCurrentRole = DataOrig %>% group_by(YearsInCurrentRole, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.YearsInCurrentRole %>% ggplot(aes(x=YearsInCurrentRole, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# YearsSinceLastPromotion - NOT important
df.YearsSinceLastPromotion = DataOrig %>% group_by(YearsSinceLastPromotion, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.YearsSinceLastPromotion %>% ggplot(aes(x=YearsSinceLastPromotion, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)


# YearsWithCurrManager - NOT important
df.YearsWithCurrManager = DataOrig %>% group_by(YearsWithCurrManager, Attrition) %>% summarise(count=n()) %>% mutate(pct=count/sum(count)) 

df.YearsWithCurrManager %>% ggplot(aes(x=YearsWithCurrManager, y=pct, fill=Attrition)) + geom_bar(stat="identity") + 
  geom_text(aes(label=paste0(round(pct*100,1),"%")), 
            position = position_stack(vjust = 0.5), size = 3)
```

```{r, include=FALSE}
# Change all character variables to factor
DataOrig$BusinessTravel.Num <- as.integer(as.factor(DataOrig$BusinessTravel))
DataOrig$Department.Num <- as.integer(as.factor(DataOrig$Department))
DataOrig$EducationField.Num <- as.integer(as.factor(DataOrig$EducationField))
DataOrig$Gender.Num <- as.integer(as.factor(DataOrig$Gender))
DataOrig$JobRole.Num <- as.integer(as.factor(DataOrig$JobRole))
DataOrig$MaritalStatus.Num <- as.integer(as.factor(DataOrig$MaritalStatus))
DataOrig$OverTime.Num <- as.integer(as.factor(DataOrig$OverTime))
```

Apply the three main factors with kNN technique to create an Attrition Prediction model
```{r}
set.seed(8)
splitPerc = .7

ImpFactors <- subset(DataOrig, select = c(Attrition, JobInvolvement, MonthlyIncome, OverTime.Num))
ImpFactors$MonthlyIncome <- scale(ImpFactors$MonthlyIncome)
#ImpFactors$Age <- scale(ImpFactors$Age)

trainIndices = sample(1:dim(ImpFactors)[1], round(splitPerc * dim(ImpFactors)[1]))
train.Attrition = ImpFactors[trainIndices,]
test.Attrition = ImpFactors[-trainIndices,]

# run knn
classifications = knn(train.Attrition[,c(2:4)], test.Attrition[,c(2:4)], train.Attrition$Attrition, prob = TRUE, k = 21)

# Confusion Matrix
confusionMatrix(table(test.Attrition$Attrition, classifications))


# Find Sensitivity and Specificity for Train Set
classifications_train = knn(train.Attrition[,c(2:4)], train.Attrition[,c(2:4)], train.Attrition$Attrition, prob = TRUE, k = 5)

confusionMatrix(table(train.Attrition$Attrition, classifications_train))
```

Loop for many k and one training / test partition
```{r}
accs = data.frame(accuracy = numeric(30), k = numeric(30))
for(i in 1:30)
{
  classifications = knn(train.Attrition[,c(2:4)], test.Attrition[,c(2:4)], train.Attrition$Attrition, prob = TRUE, k = i)
  table(test.Attrition$Attrition, classifications)
  CM = confusionMatrix(table(test.Attrition$Attrition, classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}
plot(accs$k,accs$accuracy, type = "l", xlab = "k")
```

```{r, include=FALSE}
# Loop for many k and the average of many training / test partition

iterations = 500
numks = 30
masterAcc = matrix(nrow = iterations, ncol = numks)
  
for(j in 1:iterations)
{
accs = data.frame(accuracy = numeric(30), k = numeric(30))
trainIndices = sample(1:dim(ImpFactors)[1], round(splitPerc * dim(ImpFactors)[1]))
train.Attrition = ImpFactors[trainIndices,]
test.Attrition = ImpFactors[-trainIndices,]
for(i in 1:numks)
{
  classifications = knn(train.Attrition[,c(2:4)], test.Attrition[,c(2:4)], train.Attrition$Attrition, prob = TRUE, k = i)
  table(classifications,test.Attrition$Attrition)
  CM = confusionMatrix(table(test.Attrition$Attrition, classifications))
  masterAcc[j,i] = CM$overall[1]
}
}
MeanAcc = colMeans(masterAcc)
plot(seq(1,numks,1),MeanAcc, type = "l")
```

Then write attrition predicted values for 'CaseStudy2CompSet No Attrition.csv' to csv File
```{r}
# Read in 'No Attrition' file
No.AttritionOrig <- read.csv("CaseStudy2CompSet No Attrition.csv", header=TRUE)

# Only include variables that used in the model
No.Attrition <- subset(No.AttritionOrig, select = c(ID, JobInvolvement, MonthlyIncome, OverTime))
# Use Scale with MonthlyIncome and change OverTime to interger value
No.Attrition$MonthlyIncome <- scale(No.Attrition$MonthlyIncome)
No.Attrition$OverTime <- as.integer(as.factor(No.Attrition$OverTime))

classifications = knn(train.Attrition[,c(2:4)], No.Attrition[,c(2:4)], train.Attrition$Attrition, k = 21, prob = TRUE)

# Cleanup the df to match example file before writing
Attrition.Prediction <- as.data.frame(classifications)
Attrition.Prediction$ID <- No.Attrition$ID
No.Attrition_pred <- merge(x = No.AttritionOrig, y = Attrition.Prediction,
                          by.x = "ID", by.y = "ID", all = TRUE)
names(No.Attrition_pred)[names(No.Attrition_pred)=="classifications"] <- "Attrition"
No.Attrition_pred <- No.Attrition_pred %>% relocate("Attrition", .after = "ID")

# Write to file
write.csv(No.Attrition_pred, "Case2PredictionsRudick Attrition.csv", row.names = FALSE, quote=FALSE)
```

### Linear Regression for Monthly Income Prediction

Find correlation coefficient with p-value for numerical variables
```{r}
Num_col <- subset(DataOrig, select = -c(Attrition, BusinessTravel,
                                    Department, EducationField,
                                    Gender, JobRole, MaritalStatus,
                                    Over18, OverTime,
                                    ID, EmployeeCount,
                                    EmployeeNumber, StandardHours
                                    ))
M <- cor(Num_col)
corrplot(M, type="upper")

# Run rcorr(as.matrix(Num_col)) for correlation coefficient with p-value
subset(DataOrig, select = c(MonthlyIncome, Age, JobLevel, TotalWorkingYears, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager)) %>%
  ggpairs()

# There are 7 columns that have statistically significant correlation (p-value < .05) with good range of correlation coefficient which are Age, JobLevel, TotalWorkingYears, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager

# When put all 7 columns to the linear regression model, only JobLevel and TotalWorkingYears are significant
modelNum <- lm(MonthlyIncome ~ Age+JobLevel+TotalWorkingYears+YearsAtCompany+YearsInCurrentRole+YearsSinceLastPromotion+YearsWithCurrManager, data = Num_col)
summary(modelNum)
```

Find correlation coefficient with p-value for numerical variables
```{r}
Categorical <- subset(DataOrig, select = c(MonthlyIncome, BusinessTravel,
                                           Department, EducationField,
                                           Gender, JobRole, MaritalStatus,
                                           OverTime))

modelCat <- lm(MonthlyIncome ~., data = Categorical)
summary(modelCat)

# After trying including every categorical variables to the MLR model, none of them makes the model better and/or makes the ,pde; statistically significant

```

The analyst decided to use JobLevel, TotalWorkingYears, YearsWithCurrManager for MLR model
```{r}
fit <- lm(MonthlyIncome ~ JobLevel+TotalWorkingYears+YearsWithCurrManager, data=DataOrig)
summary(fit)

# Check MLR assumptions 
par(mfrow=c(2,2))
# Check Multiple Linear Regression Asumption
plot(fit$fitted.values, fit$residuals, main = "Plot of Residuals vs. Fitted Values")

plot(DataOrig$JobLevel, fit$residuals, main = "Plot of Residuals vs. Fitted Values")

hist(fit$residuals)
#ggplot(data = fit, aes(x = fit$residuals)) +
#  geom_histogram() +
#  labs(title = 'Histogram of Residuals', x = 'Residuals', y = 'Frequency')

qqnorm(fit$residuals,  ylab = "Residual Quantiles")

# Find model's RMSE 
preds = predict(fit, newdata = DataOrig)

RMSE = data.frame(Observed = DataOrig$MonthlyIncome,
                  Predicted = preds)
RMSE$Residual = RMSE$Observed - RMSE$Predicted
RMSE$SquaredResidual = RMSE$Residual^2

mean(sqrt(RMSE$SquaredResidual)) #1,035.798

```

Cross Validation by creating Train and Test dataset from the original dataset
```{r}
# Separated dataset into Train and Test dataset
set.seed(8)
splitPerc = .7

MLR_trainIndices = sample(1:dim(DataOrig)[1], round(splitPerc * dim(DataOrig)[1]))
train.MonthlyIncome = DataOrig[MLR_trainIndices,]
test.MonthlyIncome = DataOrig[-MLR_trainIndices,]

# Created MLR model from Train dataset
model_fit <- lm(MonthlyIncome ~ JobLevel+TotalWorkingYears+YearsWithCurrManager, data=train.MonthlyIncome)
summary(model_fit)

# Validated model using Test dataset
model_preds = predict(model_fit, newdata = test.MonthlyIncome)

test.RMSE = data.frame(Observed = test.MonthlyIncome$MonthlyIncome,
                       Predicted = model_preds)
test.RMSE$Residual = test.RMSE$Observed - test.RMSE$Predicted
test.RMSE$SquaredResidual = test.RMSE$Residual^2

mean(sqrt(test.RMSE$SquaredResidual)) #1,091.855


# Find RMSE for Test data
model_preds_train = predict(model_fit, newdata = train.MonthlyIncome)

train.RMSE = data.frame(Observed = train.MonthlyIncome$MonthlyIncome,
                       Predicted = model_preds_train)
train.RMSE$Residual = train.RMSE$Observed - train.RMSE$Predicted
train.RMSE$SquaredResidual = train.RMSE$Residual^2

mean(sqrt(train.RMSE$SquaredResidual)) #1,015.037
```

Write salary predicted values for 'CaseStudy2CompSet No Salary.xlsx' to csv File
```{r}
# Read in 'No Attrition' file
No.SalaryOrig <- read_excel("CaseStudy2CompSet No Salary.xlsx")

# Only include variables that used in the model
No.Salary <- subset(No.SalaryOrig, select = c(ID, JobLevel, TotalWorkingYears, YearsWithCurrManager))

model_preds_csv = predict(model_fit, newdata = No.Salary)

# Cleanup the df to match example file before writing
MonthlyIncome.Prediction <- as.data.frame(model_preds_csv)
MonthlyIncome.Prediction$ID <- No.Salary$ID
No.Salary_pred <- merge(x = No.SalaryOrig, y = MonthlyIncome.Prediction,
                          by.x = "ID", by.y = "ID", all = TRUE)
names(No.Salary_pred)[names(No.Salary_pred)=="model_preds_csv"] <- "MonthlyIncome"
No.Salary_pred <- No.Salary_pred %>% relocate("MonthlyIncome", .after = "ID")

# Write to file
write.csv(No.Salary_pred, "Case2PredictionsRudick Salary.csv", row.names = FALSE, quote=FALSE)
```

Used SetepWise Selection to help confirm variable prediction
```{r message=FALSE, warning=FALSE}
selectedVariables <- subset(DataOrig, select = c(Attrition, MonthlyIncome, BusinessTravel, Department, EducationField,
                                    Gender, JobRole, MaritalStatus,
                                    OverTime, Age, JobLevel, TotalWorkingYears, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, YearsWithCurrManager
                                    ))
df <- subset(DataOrig, select = -c(Attrition,
                                    Over18,
                                    ID, EmployeeCount,
                                    EmployeeNumber, StandardHours
                                    ))
  
# Fit the full model 
full.model <- lm(MonthlyIncome ~., data = df)
# Stepwise regression model
step.model <- stepAIC(full.model, direction = "both", 
                      trace = FALSE)
summary(step.model)

# Set seed for reproducibility
set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model <- train(MonthlyIncome ~., data = df,
                    method = "leapSeq", 
                    tuneGrid = data.frame(nvmax = 1:5),
                    trControl = train.control
                    )
step.model$results

step.model$bestTune

summary(step.model$finalModel)


# Stepwise Selection includes JobLevel, TotalWorkingYears, and JobRole but only 2 JobRole out of 8 have statistically significant when include in the MLR
coef(step.model$finalModel, 4)
```
